Robust Estimation, Prediction and Control with \\ Linear Dynamics and Arbitrary Costs

### 2020-01-25 main.tex
Updates from Overleaf

Since $R$ is arbitrary, solving the optimal -- not to mention the robust -- control objective in the original continuous control space $(\Real q)^\Natural$ is intractable.
In this work, we choose to approximate the solution by turning the continuous optimisation problem \eqref{eq:robust-control} into a discrete one, by optimising over a finite set of \emph{actions} $\cA$. Each action $a\in\cA$ corresponds to the selection of a low-level controller $\pi_a$, that we take affine:
\[u(t) = \pi_a(x(t)) \eqdef -K_a x(t) + u_a\]

@@

\paragraph{Robust planning}
Robust optimization has been studied in the context of finite Markov Decision Processes with uncertain parameters by Iyengar (2005), Nilim and El Ghaoui (2005) and Wiesemann et al. (2013). They show that the main results of Dynamic Programming can be extended to their robust counterparts only when the dynamics ambiguity set verifies a certain rectangularity property, that amounts to an independence assumption between the uncertainty sets of each transition and is related to time-varying uncertainty.
They do not consider continuous states, nor handle time-invariant uncertainty. We do so by not relying on a dynamic programming evaluation of the robust cost-to-go, but on monte-carlo estimates along a full trajectory.

@@

\subsection{Tighter conversion form ellipsoid to polytope}
\label{sec:tight-polytope}
\begin{lemma}[Confidence polytope]
\label{lem:tight_polytope}
We can enclose the confidence ellipsoid obtained in $\eqref{eq:confidence-ellipsoid}$ within a polytope $C_\delta$:
\begin{equation}
     \cC_\delta = \left\{ A_{1}+\sum_{i=1}^{2^d}\lambda_{i}\Delta A_{i}: \lambda\in[0, 1]^{2^d},  \sum_{i=1}^{2^d}\lambda_{i}=1\right\}.
\end{equation}
with 
\begin{align*}
    &h_k \text{ is the }k^\text{th}\text{ element of }\{-1,1\}^d\text{ for } k\in[2^d],\\
    &G_{N,\lambda}^{-1} = PDP^{-1}, \quad \Delta\theta_k = C_{N,\lambda,\delta}^{1/2} P^{-1}D^{-1/2} h_k, \\
    &A_0 = A + \theta_{N,\lambda}^\transp\Phi, \quad \Delta A_k = \Delta\theta_k^\transp\Phi.
\end{align*}
This conversion is illustrated in \autoref{fig:ellipsoid_to_polytope}.
\end{lemma}

\begin{proof}
The ellipsoid in \eqref{eq:confidence-ellipsoid} is described by:
\begin{align*}
    \theta\in\cC_\delta &\implies
    (\theta-\theta_{N,\lambda})^\transp G_{N,\lambda}^{-1}(\theta-\theta_{N,\lambda}) \leq C_{N,\lambda,\delta}\\
    &\implies (\theta'-\theta'_{N,\lambda})^\transp D (\theta'-\theta'_{N,\lambda}) \leq C_{N,\lambda,\delta}\\
    &\implies \sum_{i=1}^d D_{i,i}(\theta'_i-\theta'_{N,\lambda,i})^2\leq C_{N,\lambda,\delta}\\
    &\implies\forall i, |\theta'_i-\theta'_{N,\lambda,i}|\leq C_{N,\lambda,\delta}^{1/2}D_{i,i}^{-1/2}
\end{align*}
This describes a $\Real^d$ box containing $\theta' = P\theta$, whose $k^\text{th}$ vertex is represented by $\theta_{N,\lambda}' + C_{N,\lambda,\delta}^{1/2}D^{-1/2} h_k$. We obtain the corresponding box on $\theta$ by transforming each vertex of the box with $P^{-1}$.
\end{proof}

>>>>

Since $R$ is arbitrary, solving the optimal -- not to mention the robust -- control objective in the original continuous control space $(\Real^q)^\Natural$ is intractable.
In this work, we choose to approximate the solution by turning the continuous optimisation problem \eqref{eq:robust-control} into a discrete one, by optimising over a finite set of \emph{actions} $\cA$. Each action $a\in\cA$ corresponds to the selection of a low-level controller $\pi_a$, that we take affine:
\[u(t) = \pi_a(x(t)) \eqdef -K_a x(t) + u_a\]

@@

\paragraph{Robust planning}
Robust optimization has been studied in the context of finite Markov Decision Processes with uncertain parameters by Iyengar (2005), Nilim and El Ghaoui (2005) and Wiesemann et al. (2013). They show that the main results of Dynamic Programming can be extended to their robust counterparts only when the dynamics ambiguity set verifies a certain rectangularity property, that amounts to an independence assumption between the uncertainty sets of each transition and is related to time-varying uncertainty.
L1/KL ball around the empirical transition.
They do not consider continuous states, nor handle time-invariant uncertainty. We do so by not relying on a dynamic programming evaluation of the robust cost-to-go, but on monte-carlo estimates along a full trajectory.

@@

\subsection{Tighter conversion form ellipsoid to polytope}
\label{sec:tight-polytope}
\begin{lemma}[Confidence polytope]
\label{lem:tight_polytope}
We can enclose the confidence ellipsoid obtained in $\eqref{eq:confidence-ellipsoid}$ within a polytope $C_\delta$:
\begin{equation}
     \cC_\delta = \left\{ A_{1}+\sum_{i=1}^{2^d}\lambda_{i}\Delta A_{i}: \lambda\in[0, 1]^{2^d},  \sum_{i=1}^{2^d}\lambda_{i}=1\right\}.
\end{equation}
with 
\begin{align*}
    &h_k \text{ is the }k^\text{th}\text{ element of }\{-1,1\}^d\text{ for } k\in[2^d],\\
    &G_{N,\lambda} = PDP^{-1}, \quad \Delta\theta_k = \beta_{N}(\delta)^{1/2} P^{-1}D^{-1/2} h_k, \\
    &A_0 = A + \theta_{N,\lambda}^\transp\Phi, \quad \Delta A_k = \Delta\theta_k^\transp\Phi.
\end{align*}
This conversion is illustrated in \autoref{fig:ellipsoid_to_polytope}.
\end{lemma}

\begin{proof}
The ellipsoid in \eqref{eq:confidence-ellipsoid} is described by:
\begin{align*}
    \theta\in\cC_\delta &\implies
    (\theta-\theta_{N,\lambda})^\transp G_{N,\lambda}(\theta-\theta_{N,\lambda}) \leq \beta_{N}(\delta)\\
    &\implies (\theta'-\theta'_{N,\lambda})^\transp D (\theta'-\theta'_{N,\lambda}) \leq \beta_{N}(\delta)\\
    &\implies \sum_{i=1}^d D_{i,i}(\theta'_i-\theta'_{N,\lambda,i})^2\leq \beta_{N}(\delta)\\
    &\implies\forall i, |\theta'_i-\theta'_{N,\lambda,i}|\leq \beta_{N}(\delta)^{1/2}D_{i,i}^{-1/2}
\end{align*}
This describes a $\Real^d$ box containing $\theta' = P\theta$, whose $k^\text{th}$ vertex is represented by $\theta_{N,\lambda}' + \beta_{N}(\delta)^{1/2}D^{-1/2} h_k$. We obtain the corresponding box on $\theta$ by transforming each vertex of the box with $P^{-1}$.
\end{proof}
